Mutual Information Plots

### 2025-11-06 mutual_information_curves.qmd
updated weighted kendall

@@

weighted_kendall <- function(rhat, k, wtype = "symmetric_salience", salient_weights = NULL, r = NULL) {
  n <- length(rhat)
  if (is.null(r)) r <- 1:n
  if (is.null(salient_weights)) salient_weights <- rep(1, k)

@@

  # Compute weighted Kendall numerator and denominator consistently
  summ <- 0
  denom <- 0
  for (i in 1:n) {
    for (j in 1:n) {
      if (i != j && w[i, j] != 0) {
        # Skip uninformative pairs (ties in either vector)
        if (r[i] == r[j] || rhat[i] == rhat[j]) next

        denom <- denom + abs(w[i, j])
        summ <- summ + w[i, j] * sign(r[i] - r[j]) * sign(rhat[i] - rhat[j])
      }
    }
  }

  # Handle case where denom = 0 (e.g., all ties)
  if (denom == 0) return(1)

@@

weighted_kendall(c(2,1,3,4), k=4, salient_weights = c(3,1,1,0))
weighted_kendall(c(2,1,3,4), k=2)

weighted_kendall(c(1,1,1,1,5,6), k = 6, r = c(1,1,1,1,5,6))
# Only works for k = # of top teams tied

@@

# Create plot
ggplot(plot_data, aes(x = x, y = y, color = type, group = type)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_x_continuous(breaks = 1:8) +
  labs(
    x = "Salient Ranks",
    y = "Value",
    color = "Series",
    title = "Comparison of MI vs Top K-teams"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "top",
    plot.title = element_text(hjust = 0.5)
  )

>>>>

##

```{r}
r <- c(1,2,3,4,5)
rhat <- c(4,1,1,1,5)
n <- length(r)
k <- sum(rhat == 1)
count <- 0
for (i in 1:(length(r)-1)){
  for (j in ((i+1):(length(r)))){
    print(sign(r[i] - r[j])*sign(rhat[i] - rhat[j]))
    count <- count + sign(r[i] - r[j])*sign(rhat[i] - rhat[j])
  }
}

count
#Correlation
count / (choose(n,2) - choose(k,2))
#Weight
((count / (choose(n,2) - choose(k,2))) + 1)/2

@@

weighted_kendall <- function(rhat, k, wtype = "symmetric_salience", salient_weights = NULL) {
  n <- length(rhat)
  #r <- 1:n
  if (is.null(salient_weights)) salient_weights <- rep(1, k)

@@

  # Compute weighted Kendall numerator and denominator consistently
  summ <- 0
  denom <- 0
  for (i in 1:(n-1)) {
    for (j in (i+1):n) {
      if (i != j) {
        sign_ij <- sign(i - j)
        sign_rhat <- sign(rhat[i] - rhat[j])
        sign_rhat <- ifelse(sign_rhat == 0, -1, sign_rhat)
        summ <- summ + w[i, j] * sign_ij * sign_rhat
        denom <- denom + w[i, j]
        #print(c(sign_r, sign_rhat, summ, denom))
      }
    }
  }

@@

weighted_kendall(c(2,1,3,4), k=4, salient_weights = c(3,1,1,0))
weighted_kendall(c(1.5,3,1.5), k=2)
weighted_kendall(c(1,3,4,2), k=3)
weighted_kendall(c(1,1,4,1), k=3)

# Define the function calls
results <- data.frame(
  rhat = rep(c("(1,1,3)", "(1,3,1)", "(3,1,1)"), each = 3),
  k = rep(1:3, times = 3),
  weighted_kendall = c(
    weighted_kendall(c(1,1,3), k=1),
    weighted_kendall(c(1,1,3), k=2),
    weighted_kendall(c(1,1,3), k=3),
    weighted_kendall(c(1,3,1), k=1),
    weighted_kendall(c(1,3,1), k=2),
    weighted_kendall(c(1,3,1), k=3),
    weighted_kendall(c(3,1,1), k=1),
    weighted_kendall(c(3,1,1), k=2),
    weighted_kendall(c(3,1,1), k=3)
  )
)

results

weighted_kendall(c(1,1,1,1,5,6), k = 3, r = c(1,1,1,1,5,6))
# Only works for k = # of top teams tied

@@

# Create plot
ggplot(plot_data, aes(x = x, y = y, color = type, group = type)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_x_continuous(breaks = 1:8) +
  labs(
    x = "Salient Ranks",
    y = "Value",
    color = "Series",
    title = "Comparison of MI vs Top K-teams"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "top",
    plot.title = element_text(hjust = 0.5)
  )
```
